---
title: "STA_445_HW1"
author: "Bianca L."
date: "2023-10-03"
output: pdf_document
---
```{r}
library(readxl)
```

## Problem 1
Create a vector of three elements (2,4,6) and name that vector `vec_a`. Create a second vector, `vec_b`, that contains (8,10,12). Add these two vectors together and name the result `vec_c`.

```{r}
vec_a <- c(2, 4, 6)
vec_b <- c(8, 10, 12)
(vec_c <- vec_a + vec_b)
```

## Problem 2
Create a vector, named `vec_d`, that contains only two elements (14,20). Add this vector to `vec_a`. What is the result and what do you think R did (look up the recycling rule using Google)? What is the warning message that R gives you?
```{r}
vec_d <- c(14, 20)
vec_a + vec_d
```
Since a has three elements and d only has 2, the first element of d is used twice.  The warning message lets us know that the vectors are not of the same length.

## Problem 3

Next add 5 to the vector vec_a. What is the result and what did R do? Why doesn't in give you a warning message similar to what you saw in the previous problem?

```{r}
5 + vec_a
```
R adds 5 to each element of a.  There is no warning since this is the case of a scalar plus a vector.

## Problem 4
Generate the vector of integers $\left\{ 1,2,\dots5\right\}$ in two different ways. 

a.  First using the `seq()` function

```{r}
seq(1, 5)
```

b. Using the `a:b` shortcut.
```{r}
1:5
```

## Problem 5
Generate the vector of even numbers $\left\{ 2,4,6,\dots,20\right\}$ 

a. Using the seq() function and 

```{r}
seq(2, 20, by=2)
```
b. Using the a:b shortcut and some subsequent algebra. 

```{r}
1:10*2
```
## Problem 6
Generate a vector of 21 elements that are evenly placed between 0 and 1 using the `seq()` command and name this vector `x`. 

```{r}
(x <- seq(0, 1, length=21))
```
## Problem 7
Generate the vector $\left\{ 2,4,8,2,4,8,2,4,8\right\}$ 
  using the `rep()` command to replicate the vector c(2,4,8). 
  
```{r}
rep(c(2,4, 8), 3)
```

## Problem 8
Generate the vector $\left\{ 2,2,2,2,4,4,4,4,8,8,8,8\right\}$
  using the `rep()` command. You might need to check the help file for rep() to see all of the options that rep() will accept. In particular, look at the optional argument `each=`.
  
```{r}
rep(c(2,4,8), each=4)
```

## Problem 9 (CH8 Prob 10)

In this problem, we will work with the matrix 

    \[ \left[\begin{array}{ccccc}
    2 & 4 & 6 & 8 & 10\\
    12 & 14 & 16 & 18 & 20\\
    22 & 24 & 26 & 28 & 30
    \end{array}\right]\]
    
a.  Create the matrix in two ways and save the resulting matrix as `M`.

i. Create the matrix using some combination of the `seq()` and `matrix()` commands.

```{r}
(M <- matrix(seq(2, 30, by=2), nrow=3, ncol=5,
       byrow = TRUE))
```

ii. Create the same matrix by some combination of multiple `seq()` commands and either the `rbind()` or `cbind()` command.

```{r}
rbind(seq(2, 10, by=2), 
      seq(12, 20, by=2), 
      seq(22, 30, by=2))
```

b. Extract the second row out of M.

```{r}
M[2,]
```
c. Extract the element in the third row and second column of M

```{r}
M[3,2]
```
## Problem 10 (CH8 Prob12)

The following code creates a `data.frame` and then has two different methods for removing the rows with `NA` values in the column `Grade`. Explain the difference between the two.

```{r, eval=FALSE}
df <- data.frame(name= c('Alice','Bob','Charlie','Daniel'),
                     Grade = c(6,8,NA,9))
    
#df[ -which(  is.na(df$Grade) ), ]

df[  which( !is.na(df$Grade) ), ]
```

The first option uses the negative sign to say in essence "not these rows".  The second option uses ! to negate na.  In other words, it says which rows are not NA and keep those rows.

## Problem 11 (CH8 Prob 14)

Create and manipulate a list.

a.  Create a list named my.test with elements
        + x = c(4,5,6,7,8,9,10)
        + y = c(34,35,41,40,45,47,51)
        + slope = 2.82
        + p.value = 0.000131
```{r}
my.test <- 
  list( x = 4:10,
        y = c(34, 35, 41, 40, 45, 47, 51),
        slope = 2.82,
        p.value = 0.000131)
```

b.  Extract the second element in the list.

```{r}
my.test[2]
```
c.  Extract the element named `p.value` from the list.

```{r}
my.test$p.value
```

## Problem 12

Download from GitHub the data file 
[Example_5.xls](https://github.com/dereksonderegger/444/raw/master/data-raw/Example_5.xls). Open it in Excel and figure out which sheet of data we should import into R. At the same time figure out how many initial rows need to be skipped. Import the data set into a data frame and show the structure of the imported data 
using the `str()` command. Make sure that your data has $n=31$ observations and the three columns are appropriately named. If you make any modifications to the data file, comment on those modifications.

```{r}
ex5 <- read_excel("Example_5.xls", sheet=2, range = "A5:C36") 

head(ex5)
str(ex5)
```
I could not get the dataset to read in directly from github.  I saved it in the same folder that the Rmd file is saved in and read it in from there.  I could 

## Problem 13

Download from GitHub the data file [Example_3.xls](https://github.com/dereksonderegger/444/raw/master/data-raw/Example_3.xls). Import the data set into a data frame and show the structure of the imported 
data using the `tail()` command which shows the last few rows of a data table. Make sure the Tesla values are `NA` where appropriate and that both  `-9999` and `NA` are imported as NA values. If you make any modifications to 
the data file, comment on those modifications.

```{r}
ex3 <- read_excel("Example_3.xls", sheet=2, range = "A1:L34", na=c("-9999", "NA" ) )

tail(ex3)
```
The last row of the data looks like its in the proper format now.
